{\bf Affine.m} -- {\it Mathematica} package for Lie algebras of finite, affine and extended affine type

### 2011-03-29 AffineLieAlgebras/CPC/article.tex
Some text on Weyl group, start of representation theory section

Consider Lie algebra $\gf$ of finite or affine type. We denote its root system by $\Delta$. Roots and weights of the algebra are represented by the data-structures \lstinline{finiteWeight} and \lstinline{affineWeight} of {\it Affine.m} package. 
Root system is determined by the set of simple roots $\alpha_{1}, \dots, \alpha_{r}$, where $r$ is the rank of the algebra. So to specify an algebra it is enough to have simple roots. The set of simple roots can be created using different methods. The coordinates of simple roots can be entered by hand, for example direct sum $sl_{2}\oplus sl_{2}$ can be constructed by the code
\begin{lstlisting}
  makeFiniteRootSystem[{{1,0}, {0,1}}]
\end{lstlisting}
For simple finite-dimensional Lie algebras there are built-in constructors
\begin{lstlisting}
  b2=makeSimpleRootSystem[B,2]

  Out[6]= finiteRootSystem[2, {finiteWeight[2, {1, -1}], finiteWeight[2, {0, 1}]}]
\end{lstlisting}

>>>>

Consider Lie algebra $\gf$ of finite or affine type. We denote its Cartan subalgebra by $\hf$ and root system by $\Delta$. Roots $\alpha_i\in \Delta$ and weights $\mu,\nu\in \hf^*$ of the algebra are represented by the data-structures \lstinline{finiteWeight} and \lstinline{affineWeight} of {\it Affine.m} package. 
Root system can be reconstructed from the set of simple roots $\alpha_{1}, \dots, \alpha_{r}$, where $r$ is the rank of the algebra, by the action of Weyl group $W_{\gf}$.

The set of simple roots can be created using different methods. The coordinates of simple roots can be entered by hand, for example direct sum $sl_{2}\oplus sl_{2}$ can be constructed by the code
\begin{lstlisting}
  makeFiniteRootSystem[{{1,0}, {0,1}}]
\end{lstlisting}
For simple finite-dimensional Lie algebras there are built-in constructors
\begin{lstlisting}
  b2=makeSimpleRootSystem[B,2]

  Out[6]= finiteRootSystem[2, 2, {finiteWeight[2, {1, -1}], finiteWeight[2, {0, 1}]}]
\end{lstlisting}

It is possible to specify direct sum of algebras:
\begin{lstlisting}
  b2+b2

  Out[6]= finiteRootSystem[4, 4, {finiteWeight[4, {1, -1, 0, 0}], finiteWeight[2, {0, 1}]}]
\end{lstlisting}

Addition, multiplication by number and scalar product are defined for roots and weights:
\begin{lstlisting}
  makeFiniteWeight[{1,1}]+2*makeFiniteWeight[{1,-1}]

  Out[7]=finiteWeight[2, {3, -1}]
\end{lstlisting}

It is possible to compute Cartan matrix or draw Dynkin diagram of the algebra or specify an algebra with Cartan matrix or Dynkin diagram. TODO

Action of Weyl group $W_{\gf}$ generated by simple reflections $s_{\alpha}, \alpha\in \Delta$ is important for computations. We denote by $C_{\gf}\subset \hf^*$ the main Weyl chamber (fundamental domain for the action of $W_{\gf}$) and its closure by $\bar C_{\gf}$. In many computations it is possible to work inside the main Weyl chamber due to Weyl symmetry.

Sometimes it can be helpful to consider Weyl group elements abstractly withouth the reference to root system. For example, there exists package for rule-based simplification of Weyl group elements based on finite automata [cite!]. We offer the function \lstinline{weylGroupElement[x__Integer][rs_?rootSystemQ]} which produces the object, representing element of Weyl group encoded by the set of reflections, that can act upon the weights. It can be used as follows:
\begin{lstlisting}
  <<simplification.m;
  (weylGroupElement @@ simplify[s[1,2,3,1,3,1,2,1,4,2,1]])[b2+b2][makeFiniteWeight[{1,1,0,0}]]

  Out[7]= finiteWeight[4, {-1, 1, 0, 0}]
\end{lstlisting}

Following functions can be used to compute objects related to Weyl group and its action on roots and weights
 \begin{itemize}
 \item   $\forall \mu\in \hf^* \exists \lambda\in \bar C_{\gf}, w\in W_{\gf}: \mu=w \lambda$, the function \lstinline{toFundamentalChamber} can be used to find such $\lambda$.
 \item Orbit of weight under the action of Weyl group is computed with the function \lstinline{orbit}, the elements of the orbit are grouped by the length of Weyl group element.
 \end{itemize}

\subsection{Representation theory}
\label{sec:repr-theory}

Weights of the irreducible highest weight representation are calculated using the function \lstinline{weightSystem}
\begin{lstlisting}
  weightSystem[b2][fundamentalWeights[b2][[1]]*2]
\end{lstlisting}
